oscar_figures

### 2021-03-16 oscar_figures.Rmd
add heptamer analysis

```{r, warning=FALSE}
ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
    geom_bar(stat="identity") +
    xlim(-50, 50) +
    ylab("Counts") + 
    xlab("Distance of 5' end from annotated start codon") +
    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
```

@@

>>>>

```{r, warning=FALSE}
ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
    geom_line() +
    geom_point() +
    xlim(-50, 50) +
    ylab("Counts") + 
    xlab("Distance of 5' end from annotated start codon") +
    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
```

@@

## Heptameric analysis

Analyse E-values (higher E value = stronger binding enrichment) for the motifs present at the exon SNP.

```{r message=FALSE, warning=FALSE}

# Read in supplementary 4 from Ray et al 2013

data <- read_tsv(paste0(here(), "/data/heptamers/41586_2013_BFnature12311_MOESM334_ESM.txt"))

for(type in c("exon", "intron")){

    if(type == "exon"){
        snp <-     "TAAAAGCATGGAT"
        healthy <- "TAAAAGGATGGAT"
    } else {
        snp <-     "GGATGGGTGGATA"
        healthy <- "GGATGGTTGGATA"
    }

    all_healthy <- c()
    all_risk <- c()

    # Make dataframe containing E scores for each relevant 7mer

    for(i in 1:(str_length(snp)-6)){
        this_snp <- str_sub(snp, i, i+6)
        this_healthy <- str_sub(healthy, i, i+6)

        all_healthy <- c(all_healthy, str_replace_all(this_healthy, "T", "U"))
        all_risk <- c(all_risk, str_replace_all(this_snp, "T", "U"))

        row1 <- data %>% filter(`7mer` == str_replace_all(this_snp, "T", "U")) %>%
            mutate(snp = T, pos = i)

        row2 <- data %>% filter(`7mer` == str_replace_all(this_healthy, "T", "U")) %>%
            mutate(snp = F, pos = i)

        if(i == 1){
            full <- bind_rows(row1, row2)
        } else {
            full <- bind_rows(full, row1, row2)
        }
    }

    # Find average change in E score for the two groups of heptamers

    full2 <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        pivot_longer(cols = contains("set")) %>%
        select(-`7mer`) %>%
        pivot_wider(names_from = "snp", values_from = "value") %>%
        mutate(abs_change = `TRUE`-`FALSE`) %>%
        filter(!is.na(`TRUE`) & !is.na(`FALSE`)) %>%
        group_by(name) %>%
        mutate(mean_abs_change = mean(abs_change)) %>%
        select(name, mean_abs_change) %>%
        unique() %>%
        arrange(mean_abs_change) %>%
        mutate(is_tdp43 = ifelse(str_detect(name, "RNCMPT00076"), "red", "grey80")) %>%
        mutate(short_name = str_sub(name, 1, 11)) %>%
        ungroup() %>% group_by(short_name) %>%
        mutate(average_delta = mean(mean_abs_change)) %>%
        select(short_name, is_tdp43, average_delta) %>%
        unique()

    # Plot ranks

    plot <- ggplot(full2, aes(x=rank(average_delta), y = average_delta,
                      fill = is_tdp43)) +
        geom_bar(stat="identity") +
        scale_fill_identity() +
        theme_minimal() +
        ggtitle(paste0("Average change in binding E value for each dataset with SNP for ", type, " SNP")) +
        xlab("") +
        ylab("Average change in E score") +
        ggpubr::theme_pubr() +
        xlab("Rank")

    print(plot)

    ggsave(paste0("C:/Users/ogw/Google Drive/UCL PhD/Year 2/unc13a/figures for paper/heptamter ranking ",
                  type, ".svg"), plot = plot)

    full_just_tdp <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        select(`7mer`, contains("RNCMPT00076"), snp, pos) %>%
        mutate(mean_E = (RNCMPT00076_e_setA + RNCMPT00076_e_setB)/2) %>%
        pivot_longer(cols = contains("RNCMP")) %>%
        mutate(ypos = ifelse(snp, -0.35, -0.27))

    plot2 <- ggplot(full_just_tdp, aes(x = pos, y = value, colour = snp,
                              label = `7mer`)) +
        geom_point(size = 3) +
        xlab("7mer") +
        ylab("E value (similar to rank?)") +
        ggeasy::easy_add_legend_title("Contains Risk SNP") +
        ggtitle(paste0("Binding E value for TDP43 for all 7x 7mers with or without risk SNP for ", type, " SNP")) +
        geom_label(aes(y = ypos), size = 2.7) +
        xlim(0.5, 7.5) +
        theme_bw() + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  +
        ggeasy::easy_remove_x_axis()

    print(plot2)

}
